### 2011-03-03 qualificacao/cap-trabrelacionados.tex
planejamento
eliminando referencias nao utilizadas no bibtex

citar tabela do madeyski com os experimentos (e colocar o trabalho dele na tabela)

@@

>>>>

Many empirical experiments have been done in order to evaluate TDD's effects on
both internal and external quality. Also, these experiments can be divided
into two categories: industry and academia. Most of experiments in industry
show results similar to what is found in this study.

When talking about external quality, Janzen \cite{janzen} demonstrated that
programmers using TDD in industry produced code that passed in up to 50\% more 
external tests than code produced by control groups not using TDD and spent less
time to fix defects. A study from George and Williams \cite{george-williams}
that produced code passed between 18\% and 50\% more in external test cases than the
code produced by groups not using TDD. The study from Edwards \cite{edwards},
with 59 students showed that TDD code has 45\% less defects.

Although many other studies relate TDD and external quality, no participants 
mentioned effects on external quality, which may indicate that TDD effects are
only in internal quality and external quality is just a side-effect.

Janzen \cite{janzen} also reported that computational complexity is much lower
in test-first code while test volume and coverage are higher. 
The same study from George and Williams also found that the code presented a
test coverage between 92\% and 98\%. Test coverage was
only mentioned (not discussed) by the participants.

A qualitative analysis in George and Williams showed that 87.5\% of the
programmers believed that TDD approach facilitated requirements 
understanding and 95.8\% believed that it reduced debugging effort. 
Regarding quality, 92\% of the developers believed that TDD yielded higher
quality code and 79\% thought it promoted simpler design.

Another study from Janzen \cite{janzen-2} with three different academic groups 
(each one using a different approach: test-first, test-last, no test) found that
the code produced by the test-first team better used object-oriented concepts,
and responsibilities were separated in thirteen different classes while the
other teams produced a more procedural code.
The test-first team also produced more code and delivered more features.
Moreover, tests produced by the test-first team had twice more assertions than
the others and covered 86\% more branches than the test-last team. Furthermore,
tested classes had 104\% lower coupling measures than untested classes and
tested methods were 43\% on average less complex than the untested ones.

Langr also \cite{langr} showed that TDD improved code quality, provided
better maintainability, and produced 33\% more tests.
Steinberg \cite{steinberg} showed that TDD code was more cohesive and less
coupled and students reported that defects were easier to fix.

Participants agree with most of this results; in their opinion, TDD improves
code quality both in implementation and design, also making it as simple as possible.
The written code is also easier to maintain as the test suite ensures the
behavior, giving freedom to programmers to refactor the code.

Higher productivity was also an effect claimed by the participants of this
study. In their opinion, although programmers do not feel productive at the
beginning, in medium terms productivity grows.

This result is also confirmed by George and
Williams who also found that although TDD might initially reduce productivity
among inexperienced programmers, in a qualitative analysis, 78\% of them thought that 
TDD improved overall programming productivity. However only 50\% of them 
believed that TDD led to less code development time.
A study from Erdogmus et al \cite{erdogmus} with 24 undergraduate students
showed that TDD increased productivity.

However, a few experiments with students show a different result. Müller and
Hagner \cite{muller} study showed that TDD has no quality and productivity 
effects. However, students noticed a better reuse in a TDD code.
Pancur \cite{pancur} showed an experiment with 38 students in which they did not
notice any quality effect or productivity improvement. In fact, students 
thought TDD was not a very effective practice.

Also, in the study from Erdogmus et al with students, no differences between
quality effects in TDD code were found. It is interesting to notice that most
experiments with students do not show results in favor of TDD. It may indicate
that students still do not have enough experience to see its effects on design.

The learning difficulty presented by the participants was also evaluated by
researchers. Mugridge \cite{mugridge} identified two main challenges in
teaching TDD over the last two years: to get students to rethink about design, 
and to really engage with this new approach. Also, it is
hard to explicitly develop students' skills in testing, design and refactoring.

Participants sometimes mentioned steps that are not suggested by TDD cycle. This
is not anomalous. Aniche and Gerosa \cite{aniche-gerosa} reported that
programmers sometimes do not follow TDD steps as described by Kent Beck \cite{tdd-by-example}.
Some mistakes identified were: to forget the refactoring step, to build complex
test scenarios, and refactor another piece of code while working on a test.
Some mistakes are frequently made by around 25\% of programmers.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
citar tabela do madeyski com os experimentos (e colocar o trabalho dele na tabela)

@@

critica aos trabalhos relacionados
